---
title: "Course Schedule"
output:
  html_document:
    toc: true
    toc_depth: 1
---

```{r, echo=FALSE, message=FALSE, results='hide', purl=FALSE}
library(tidyverse)
source("functions.R")
yamls=yaml_dir()

yamls=yamls[order(sapply(yamls,function(x) x$week))]
weeks=read_csv("weeks.csv")
```

Course content is designed to be flexible to accommodate student interest and abilities.  The order and timing of course topics may change as the semester progresses.  

**Readings**

In the table below, R4DS refers to [_R for Data Science_ by Garrett Grolemund and Hadley Wickham](http://r4ds.had.co.nz) and, while GCR refers to [_Geocomputation with R_ by Robin Lovelace, Jakub Nowosad, and Jannes Muenchow (CRC Press, 2018)](https://geocompr.robinlovelace.net/).

**Homework**
Homeworks are due at 5pm on the Friday of the week specified below.


## Schedule

```{r table,results='asis',echo=F}
# This chunk prints an index of all the task titles along with a preview of the sub-tasks 
cat("| Week | Date | Presentation | Topic / Tasks  |  Case Study | [Package Present](PackageIntro.html) |\n")
cat("|:--|:--|:-:|:-:|:-:|:---|:-:|\n") 
for(i in weeks$week){
  #print(i)
  # Find items for that week
yamlweeks=sapply(yamls,function(x) x$week)
yamltype=sapply(yamls,function(x) x$type)

wtasks=yamls[yamlweeks==i & yamltype=="Task"]
wcs=yamls[yamlweeks==i & yamltype=="Case Study"]

wtasks2=ifelse(length(wtasks)==1, 
               paste0("[",wtasks[[1]]$title,"](",wtasks[[1]]$url,")"), 
               " - ")

wcs2=ifelse(length(wcs)==1, 
               paste0("[",wcs[[1]]$title,"](",wcs[[1]]$url,")"), 
               " - ")

wpresent=ifelse(length(wtasks)==1,ifelse(!is.null(wtasks[[1]]$presentation), 
               paste0( "[<i class='fas fa-desktop'>    </i>](presentations/",
                       wtasks[[1]]$presentation,"){target='_blank'}"), 
               ""),"")
#cs_read=sapply(wcs,function(x) x$reading)
#task_read=sapply(wtasks,function(x) x$reading)
#wreadings=paste0("| ",paste(cs_read,task_read,collapse="; "),"| ")

  cat(paste(
    "| ",weeks$week[i],
    "| ",weeks$date[i],
#    wreadings,
    "| ",wpresent,
    "| ",wtasks2,
    "| ",wcs2,
    "| ",ifelse(is.na(weeks$package[i]),"",weeks$package[i]),
    "|\n "
    ))
}
```
