---
title:  Project Proposal
week: 7
type: Task
subtitle: 
presentation: PS_07_help.html
reading:
  - Chapters [3-4 in GCR](https://geocompr.robinlovelace.net/) 
tasks:
  - Complete project proposal and upload .Rmd and .md to Github
---

# Reading

```{r reading,results='asis',echo=F}
md_bullet(rmarkdown::metadata$reading)
```

### Project Proposal

The project proposal will be about 1 page in length and include the following:
  
1.  Introduction to problem/question
2.  Links to inspiring examples:  Include links and screenshots of a few (~3-5) example graphics found on the internet that convey what you want to do.  Include a few sentences about why you selected each link.
2.  Proposed data sources: Be as specific as possible.
3.  Proposed methods: List the approach you will use. Don't simply list packages, tell me what you will do with them.
4.  Expected results: Describe what you want to produce (graphics, analyses, etc.)

Finding good data takes time, and can take longer than the time to tidy your data.  This task could easily take 3-6 hours to find the data you need for your project.  After you find good data sources make sure to complete the remaining tasks.

```{r echo=F, results='hide'}
file.copy("Project_ProposalTemplate.Rmd","scripts/Project_ProposalTemplate.Rmd")
```

Use [this template (which is an R Markdown File) for your project proposal](scripts/Project_ProposalTemplate.Rmd).  You can see what this file looks like 'after rendering' [here](Project_ProposalTemplate.html).
