---
title:  Project First Draft
subtitle:  Submit the first draft of your project for peer review
week: 11
type: Task
presentation: PS_11_ParallelProcessing.html
reading:
  - Documentation for [RMarkdown Websites](https://rmarkdown.rstudio.com/rmarkdown_websites.htm)
tasks:
   - Commit your first draft of your project to GitHub
---

```{r, echo=FALSE, message=FALSE, results='hide', purl=FALSE}
source("functions.R")
source("knitr_header.R")
```

`r #presframe()`

# Reading

```{r reading,results='asis',echo=F}
md_bullet(rmarkdown::metadata$reading)
```

# Tasks

```{r reading,results='asis',echo=F}
md_bullet(rmarkdown::metadata$tasks)
```

### First Draft

The first draft of your project will be assessed by your peers in GitHub. The objectives of the peer evaluation are:

* Expose you to the work of your peers so you know what others are working on
* Provide an opportunity to share your knowledge to improve their project

You should use the project website template (or similar) to generate a html version of your project report. If your project requires any data not available in public repositories, you should put it in a folder called `/data` in your project's home directory and then import it into R with `read.csv('data/filname.csv')` or similar so that anyone with a copy of the repository can re-create the HTML output.

## Required components of first draft

1) **Introduction**  [~ 200 words]: Clearly stated background and questions / hypotheses / problems being addressed. Sets up the analysis in an interesting and compelling way.
2) **Data**: Script downloads at least one dataset automatically through the internet or loads the data from the `data/` folder.  This could use a direct download (e.g. download.file()) or an API (e.g. anything from ROpenSci).
3) **Figure**: The HTML file includes at least one figure of the data.
2) **Reproducibility**: The .Rmd should generate the HTML output when "Build Website" is clicked.

### Confirming 'reproducibility'

After pushing the files to GitHub, try downloading it as a zip file, opening in RStudio, and clicking build website - it should work.

## Common issues

1) Importing data from somewhere on your computer.  You should not have any commands such as `read.csv("~/projects/inputdata.csv")` that read any data from your computer other than the `data/` folder in your repository.