---
title: "Getting Help!"
subtitle: Learning more about finding help
reading:
   - How to [write a reproducible example](http://adv-r.had.co.nz/Reproducibility.html)
presentation:
   - day_12_help.html
tasks:
   - Learn how to read R help files effectively
   - Learn how to search for help
   - Learn how to create a Minimum Working Example (MWE)
   - Debug existing code
   - Save your reprex to your course repository as an html file using Export -> "Save As Webpage" in the RStudio "Viewer" Tab.
---

```{r, echo=FALSE, message=FALSE, results='hide', purl=FALSE}
source("functions.R")
source("knitr_header.R")
```

`r presframe()`

# Reading

```{r reading,results='asis',echo=F,purl=F}
md_bullet(rmarkdown::metadata$reading)
```

# Tasks

```{r tasks,results='asis',echo=F,purl=F}
md_bullet(rmarkdown::metadata$tasks)
```

## Download

`r output_table()`

## Libraries

```{r message=F,warning=FALSE}
library(tidyverse)
library(reprex)
library(sf)

library(spData)
data(world)
```

## Your problem

You want to make a figure illustrating the distribution of GDP per capita for all countries within each continent using the `world` data in the `spData` package.  Your desired figure looks something like the following:

```{r, echo=F, purl=F, warning=F}
library(spData)
library(ggplot2)
data(world)

ggplot(world,aes(x=gdpPercap, fill=continent))+
   geom_density(alpha=0.5,color=F)+
   labs(x="GDP Per Capita",y="Density")+
   theme(legend.position = "bottom")
```

You have started working on the figure but can't seem to make it work like you want.  Here is your current version of the code (and the resulting figure):

```{r, warning=F}
ggplot(world,aes(x=gdpPercap, y=continent, color=continent))+
   geom_density(alpha=0.5,color=F)
```

You want to ask for help and so you know that you need to make a reproducible example.  Starting with the code above, make the required edits so you can use `reprex()` to generate a nicely formatted example that you could email or post to a forum to ask for help.  See the [reading](https://reprex.tidyverse.org/) for more help. Note: you do _not_ need to recreate the figure above, only to use `reprex()` to illustrate your question.

If you have extra time, try to fix the code above to make the first figure.

