---
title: "Getting Help!"
subtitle: Learning more about finding help
reading:
   - How to [write a reproducible example](http://adv-r.had.co.nz/Reproducibility.html)
presentation:
   - day_12_help.html
tasks:
   - Learn how to read R help files effectively
   - Learn how to search for help
   - Learn how to create a Minimum Working Example (MWE)
   - Debug existing code
---

```{r, echo=FALSE, message=FALSE, results='hide', purl=FALSE}
source("functions.R")
source("knitr_header.R")
```

`r presframe()`

## Download

`r output_table()`

## Libraries

```{r message=F,warning=FALSE}
library(tidyverse)
library(reprex)
library(spData)
library(sf)
```

## Your problem

You want to make a figure illustrating the distribution of GDP per capital for all countries within each continent.  Your desired figure looks something like the following:

```{r, echo=F, purl=F}
library(spData)
library(ggplot2)
data(world)

ggplot(world,aes(x=gdpPercap, fill=continent))+
   geom_density(alpha=0.5,color=F)+
   labs(x="GDP Per Capita",y="Density")
```

You want to ask for help and so you know that you need to make a reproducible example.  Starting with the code below, use `reprex()` to generate a nicely formatted example that you could email or post to a forum to ask for help.  See the [reading](https://reprex.tidyverse.org/) for more help.

```{r}
ggplot(world,aes(x=gdpPercap, y=continent, color=continent))+
   geom_density(alpha=0.5,color=F)
```




