---
title: "Dynamic Visualization"
---
```{r setup, echo=F, results='hide'}
knitr::opts_knit$set(root.dir = NULL)
```
# HTML Visualization


## DataTables

[DataTables](http://rstudio.github.io/DT/) display R data frames as interactive HTML tables (with filtering, pagination, sorting, and search).  This is a great way to make your raw data browsable without using too much space.

```{r}
library(widgetframe)
library(DT)
datatable(iris, options = list(pageLength = 5))
```

## rbokeh

Interface to the [Bokeh](http://hafen.github.io/rbokeh) library for making interactive graphics.

```{r, warning=F, message=F}
library(rbokeh)
figure(width = 400, height=400) %>%
  ly_points(Sepal.Length, Sepal.Width, data = iris,
    color = Species, glyph = Species,
    hover = list(Sepal.Length, Sepal.Width))
```


## Leaflet

[Leaflet](http://rstudio.github.io/leaflet/) is a really powerful JavaScript library for creating dynamic maps that support panning and zooming along with various annotations like markers, polygons, and popups.  

```{r, warning=F, message=F}
library(leaflet)
library(ggmap)
geocode("Buffalo, NY")
m <- leaflet() %>% setView(lng = -78.87837, lat = 42.88645, zoom = 12) %>% 
  addTiles()
frameWidget(m,height =500)
```

## dygraphs
Provides rich facilities for charting time-series data in R, including highly configurable series- and axis-display and interactive features like zoom/pan and series/point highlighting.
    
```{r, warning=F}
library(dygraphs)
dygraph(nhtemp, main = "New Haven Temperatures",height = 100) %>% 
  dyRangeSelector(dateWindow = c("1920-01-01", "1960-01-01"))%>%
  frameWidget(height =500)

```


## rthreejs

Create interactive 3D scatter plots, network plots, and globes using the ['three.js' visualization library](https://threejs.org).
    
```{r, message=F, results=F}
#devtools::install_github("bwlewis/rthreejs")
library(threejs)
z <- seq(-10, 10, 0.1)
x <- cos(z)
y <- sin(z)
scatterplot3js(x, y, z, color=rainbow(length(z)))
```

## networkD3

Creates 'D3' 'JavaScript' network, tree, dendrogram, and Sankey graphs from 'R'.
    
```{r, message=F, results=F}
library(igraph)
library(networkD3)
```


## Load example network
This loads an example social network of friendships between 34 members of a karate club at a US university in the 1970s. See W. W. Zachary, An information flow model for conflict and fission in small groups, Journal of Anthropological Research 33, 452-473 (1977).

```{r}
karate <- make_graph("Zachary")
wc <- cluster_walktrap(karate)
members <- membership(wc)

# Convert to object suitable for networkD3
karate_d3 <- igraph_to_networkD3(karate, group = members)
```

## Force directed network plot

```{r}
forceNetwork(Links = karate_d3$links, Nodes = karate_d3$nodes,
             Source = 'source', Target = 'target', NodeID = 'name',
             Group = 'group')%>%
  frameWidget(height =500)
```


## Sankey Network graph

Sankey diagrams are flow diagrams in which the width of the arrows is shown proportionally to the flow quantity.

```{r}
# Load energy projection data
library(jsonlite)
URL <- paste0(
        "https://cdn.rawgit.com/christophergandrud/networkD3/",
        "master/JSONdata/energy.json")
Energy <- fromJSON(URL)
```

---

```{r}
sankeyNetwork(Links = Energy$links, Nodes = Energy$nodes, Source = "source",
             Target = "target", Value = "value", NodeID = "name",
             units = "TWh", fontSize = 12, nodeWidth = 30)%>%
  frameWidget(height =500)
```

## Radial Network
```{r}
URL <- paste0(
        "https://cdn.rawgit.com/christophergandrud/networkD3/",
        "master/JSONdata//flare.json")
## Convert to list format
Flare <- jsonlite::fromJSON(URL, simplifyDataFrame = FALSE)
# Use subset of data for more readable diagram
Flare$children = Flare$children[1:3]
```

---

```{r}
radialNetwork(List = Flare, fontSize = 10, opacity = 0.9, height = 400, width=400)%>%
  frameWidget(height =500)
```

## Diagonal Network
```{r}
diagonalNetwork(List = Flare, fontSize = 10, opacity = 0.9, height = 400, width=400)%>%
  frameWidget(height =500)
```


## rglwidget

RGL provides 3D interactive graphics, including functions modelled on base graphics (`plot3d()`, etc.) as well as functions for constructing representations of geometric objects (`cube3d()`, etc.).  You may need to install [XQuartz](https://www.xquartz.org/).

```{r, message=F}
library(rgl)
library(rglwidget)
library(htmltools)

theta <- seq(0, 6*pi, len=100)
xyz <- cbind(sin(theta), cos(theta), theta)
lineid <- plot3d(xyz, type="l", alpha = 1:0, 
                 lwd = 5, col = "blue")["data"]

```

## Animate an interactive 3D process
```{r}
browsable(tagList(
  rglwidget(elementId = "example", width = 500, height = 400,
            controllers = "player"),
  playwidget("example", 
             ageControl(births = theta, ages = c(0, 0, 1),
                        objids = lineid, alpha = c(0, 1, 0)),
                        start = 1, stop = 6*pi, step = 0.1, 
                        rate = 6,elementId = "player")
))%>%save_html(file="rgl.html")
```
---
<iframe class='embed-responsive-item' src='rgl.html'  width=100% height=700px allowfullscreen></iframe>

