---
title: Raster Data
date: 2018-10-04 
subtitle: Working with Raster Data
reading:
   - The [Spatial Features Package (sf)](https://r-spatial.github.io/sf/){target='blank'}
tasks:
   - Reproject spatial data using `st_transform()`
   - Intersect a vector and raster layer
   - Save your script as a .R or .Rmd in your course repository
---


```{r setup, include=FALSE, purl=F}
source("functions.R")
source("knitr_header.R")
```

# Reading

```{r reading,results='asis',echo=F,purl=F}
md_bullet(rmarkdown::metadata$reading)
```

# Background

# Objective
> 1) Generate a polygon that includes all land in NY that is within 10km of the Canadian border and 2) calculate it's area in km^2.  How much land will you need to defend from the Canadians?


# Tasks

```{r tasks,results='asis',echo=F, purl=F}
md_bullet(rmarkdown::metadata$tasks)
```

[<i class="fa fa-file-code-o fa-1x" aria-hidden="true"></i> Download starter R script (if desired)](`r output_nocomment`){target="_blank"}

<div class="well">
<button data-toggle="collapse" class="btn btn-primary btn-sm round" data-target="#demo1">Show Hints</button>
<div id="demo1" class="collapse">
The details below describe one possible approach.

## Libraries
You will need to load the following packages
```{r warning=FALSE, message=FALSE}
library(raster)
library(spData)
library(tidyverse)
```

Loading the `spData()` package returns a warning: `To access larger datasets in this package, install the spDataLarge package...`.  This is not required for this course - you can use the standard lower resolution files.

## Data
```{r  message=F}
#load 'world' data from spData package
data(world)  
```

## Steps
1. `world` dataset
    1. transform to the albers equal area projection:
    ```{r}
    albers="+proj=aea +lat_1=29.5 +lat_2=45.5 +lat_0=37.5 +lon_0=-96 +x_0=0 +y_0=0 +ellps=GRS80 +datum=NAD83 +units=m +no_defs"
    ```
    2. use `st_set_geometry()` to specify that the `geom` column contains the `geometry`.  This will also rename the column from `geom` to `geometry` to make it easier to use `ggplot()`
    3. filter the world dataset to include only `name_long=="Canada"` 
    4. buffer canada to 10km (10000m)
2. WorldClim object 
    1. Use `tmax=getData(...)` to download the WorldClim maximum temperature dataset at the lowest resolution (10 degrees). You will have to replace the `...` with the correct parameters to specify the `name="worldclim", var="tmax", res=10`.
    2. Inspect the new `tmax` object.  How many layers does it have?  What do these represent?
    3. Calculate the mean annual maximum temperature using `mean()`
4. 

</div>
</div>

Your final result should look something like this:

```{r purl=F, echo=F, message=FALSE, warning=FALSE}
world_sp <- as(world,"Spatial")

tmax <- getData(name = "worldclim", var="tmax", res=10)
gain(tmax) <- 0.1
tmax_mean <- mean(tmax)
names(tmax_mean) <- "Tmax"

world_clim <- tmax_mean %>%
  raster::extract(y=world_sp, fun=mean, na.rm=T, small=T, sp=T)%>%
  st_as_sf()

ggplot(world_clim,aes(fill=Tmax))+
  geom_sf()+
  scale_fill_viridis_c()
```

```{r}
world_clim%>%
  group_by(continent)%>%
  arrange(Tmax)%>%
  top_n(n=1,wt=Tmax)%>%
  select(name_long,continent,Tmax)%>%
  knitr::kable()
```


<div class="extraswell">
<button data-toggle="collapse" class="btn btn-link" data-target="#extras">
Extra time? Try these extra activities...
</button>
<div id="extras" class="collapse">

Build a leaflet map of the same dataset.

```{r, purl=F, echo=F}
library(leaflet)
library(htmlwidgets)
l=ny_border%>%
  st_transform("+proj=longlat +datum=WGS84")%>%
  leaflet() %>%
        addTiles() %>%
        addPolygons(color = "#444444", weight = 1, fillOpacity = 0.5,
                    fillColor = "red",
                    popup = paste("Danger Zone"))
f <-"CS05_leaflet.html"
saveWidget(l,file.path(normalizePath(dirname(f)),basename(f)),libdir="externals",selfcontained = T)
```

<iframe id="test"  style=" height:400px; width:100%;" scrolling="no"  frameborder="0" src="CS05_leaflet.html"></iframe>


</div>
</div>