---
title:  Dynamic Visualization
subtitle:  Interactive web-based visualizations
week: 12
type: Task
presentation: PS_12.html
reading:
  - Chapter [28 in R4DS](http://r4ds.had.co.nz)
tasks:
   - Browse the [Leaflet website](http://rstudio.github.io/leaflet/) and take notes in your readme.md about potential uses in your project. What data could you use?  How would you display it?
   - Browse the [HTML Widgets page](http://gallery.htmlwidgets.org/) for many more examples. Take notes in your readme.md about potential uses in your project.
---

```{r, echo=FALSE, message=FALSE, results='hide', purl=FALSE}
source("functions.R")
source("knitr_header.R")
```

`r presframe()`

# Reading

```{r reading,results='asis',echo=F}
md_bullet(rmarkdown::metadata$reading)
```

In this session we will explore several ways to generate dynamic and interactive data displays.  These include making maps and graphs that you can pan/zoom, select features for more information, and interact with in other ways.  The most common output format is HTML, which can easily be embedded in a website (such as your final project!).

```{r cache=F, message=F,warning=FALSE}
library(dplyr)
library(ggplot2)
library(ggmap)
library(htmlwidgets)
library(widgetframe)
```

If you don't have the packages above, install them in the package manager or by running `install.packages("widgetframe")`. 


