---
title: "Spatial data"
---

```{r, echo=F,message=F,warning=F}
library(methods)
library(rgdal)
library(rgeos)
library(sp)
library(sf)
library(tidyverse)
library(spData)
```

---

# Course Logistics Reminder

## Project Proposal

Start thinking about:

* Question you want to answer
* Problem you want to solve


## {data-background-iframe="https://cran.r-project.org/web/views/"}

# Working with Spatial Data in R

## Available Packages

* `sp`  First major spatial data package/format
* `rgdal` reading and writing spatial data
* `rgeos` Interface to open-source geometry engine (GEOS)
* `sf` Spatial Features in the 'tidyverse'
* `raster` gridded data (like satellite imagery)
* and a few others...

---

## What is a Spatial Feature (sf)?

Typically an object in the real world, such as a building or a tree. 

Features could include:

* a forest (polygon)
* a tree in the forest (point or polygon)
* a branch on the tree (line?)
* a complete image (multipoint, polygon, raster)
* a satellite image pixel of that forest (point or polgyon or raster)

---

```{r, echo=F}
px <- c(5, 7, 8, 9, 8, 7, 6) 
py <- c(7, 3, 4, 8, 9, 15, 14) 
plot(px, py, type="n", axes=F, xlab = '', ylab = '') 
polygon(px, py, col = "khaki1")
points(c(6, 9, 8, 8.5), c(9, 14, 8, 9), pch=20, col = "peachpuff4", lwd = 3)
lines(c(5, 6, 7, 8), c(5, 6,10, 11), col = "steelblue1",  lwd = 3)
lines(c(8, 9), c(14, 12), col = "dark green", lwd = 3)
```

---

## Spatial Features

What information do we need to store in order to define points, lines, polygons in geographic space?

 - lat/lon coordinates
 - projection
 - what type (point/line/poly)
 - if polygon, is it a hole or not
 - attribute data
 - ... ?

---

## Geometry

Features have a _geometry_ describing _where_ on Earth the
feature is located, and they have attributes, which describe other
properties. 

A Tree: 

* delineation of its crown
* its stem
* point indicating its centre

Attributes:

* species 
* height
* diameter at breast height at a particular date
* ...

---
## Spatial Feature Standard

"_A simple feature is defined by the OpenGIS Abstract specification to have both spatial and non-spatial attributes. Spatial attributes are geometry valued, and simple features are based on 2D geometry with linear interpolation between vertices._"  

---

## Dimensions

All geometries composed of points. Points are coordinates in a
2-, 3- or 4-dimensional space.  All points in a geometry have the
same dimensionality. In addition to X and Y coordinates, there are
two optional additional dimensions:

* **Z** coordinate denoting altitude
* **M** coordinate (rarely used), denoting some _measure_ that is associated with the point, rather than with the feature as a whole (in which case it would be a feature attribute); examples could be time of measurement, or measurement error.

---
## Dimensions

The four possible cases then are:

1. 2D (XY): x and y, easting and northing, or longitude and latitude
2. 3D (XYZ): three-dimensional points
3. 3D (XYM): three-dimensional points where 3rd is some attribute space
4. 4D (XYZM): four-dimensional points as XYZM (the third axis is Z, fourth M)

---
## Seven common Simple Feature (SF) geometry types

| type | description |
| ---- | ----------- |
| `POINT` | a single point |
| `LINESTRING` | sequence of points connected by lines |
| `POLYGON` | sequence of points form a closed ring |
| `MULTIPOINT` | set of points |
| `MULTILINESTRING` | set of linestrings |
| `MULTIPOLYGON` | set of polygons |
| `GEOMETRYCOLLECTION` | set of geometries |

Some formats only include these (e.g. [GeoJSON](https://tools.ietf.org/html/rfc7946))

---

## Uncommon Geometry Types 
10 more geometries 10 are rare:

* `CIRCULARSTRING`
* `COMPOUNDCURVE` 
* `CURVEPOLYGON` 
* `MULTICURVE` 
* `MULTISURFACE` 
* `CURVE` 
* `SURFACE` 
* `POLYHEDRALSURFACE` 
* `TIN` 
* `TRIANGLE`

---
## Coordinate reference system

SFs can only be placed on the Earth's surface when their coordinate
reference system (CRS) is known; this may be an elipsoidal CRS such as WGS84, a projected, two-dimensional (Cartesian) CRS such as a UTM zone or Web Mercator, or a CRS
in three-dimensions or [including time](http://www.faculty.jacobs-university.de/pbaumann/iu-bremen.de_pbaumann/Papers/acmgis-2012_crs-nts.pdf). Similarly, M-coordinates need an attribute reference
system, e.g. a [measurement unit](https://CRAN.R-project.org/package=units).


---

There are currently two main approaches in R to handle geographic vector data. 

---

# `sp` package

---

First package to provide classes and methods for spatial data types in R is called [`sp`](https://cran.r-project.org/package=sp). Provides classes and methods to create _points_, _lines_, _polygons_, and _grids_ and to operate on them. 

About 350 of the spatial analysis packages use the spatial data types that are implemented in `sp` i.e. they "depend" on the `sp` package and many more are indirectly dependent.

---

Many spatial R packages still depends on the **sp** package, therefore it is important to know how to convert **sp** to and from **sf** objects

---

Foundational structure for any spatial object in `sp` is the `Spatial` class. It has two "slots" ([new-style S4 class objects in R have pre-defined components called slots](http://stackoverflow.com/a/4714080)):

* a __bounding box__ 
* a __CRS class object__ to define the Coordinate Reference System 

Basic structure extended depending on the characteristics of the spatial object (point, line, polygon).

--- 

##  I. Create geometric objects (topology)

* __Points__ most basic spatial data objects. Typically from a two-column matrix/dataframe with a column for latitude and one for longitude.  
*  __Lines__ are generated out of `Line` objects. A `Line` object is a collection of 2D coordinates and is generated out of a two-column matrix or a dataframe with a column for latitude and one for longitude. A `Lines` object is a __list__ of one or more `Line` objects.
* __Polygons__ are generated out of `Polygon` objects. A `Polygon` object is a collection of 2D coordinates with equal first and last coordinates and is generated out of a two-column matrix or a dataframe with a column for latitude and one for longitude. A `Polygons` object is a __list__ of one or more `Polygon` objects, for example islands belonging to the same country.

---

A very simple `Line` object:
```{r}
ln <- Line(matrix(runif(6), ncol=2))
str(ln)
ln@coords
```

---

A very simple `Lines` object:
```{r}
lns <- Lines(list(ln,ln), ID = "a") # this contains just two (identical) Lines!
str(lns)
```

---

## II. Create spatial objects `Spatial*` object (`*` stands for Points, Lines, or Polygons). 

Adds bounding box and the slot for the Coordinate Reference System or CRS. `SpatialPoints` can be directly generated out of the coordinates.  `SpatialLines` and `SpatialPolygons` objects are generated using lists of `Lines` or `Polygons` objects respectively (more below).

---

See here for how to create a `SpatialLines` object:
```{r}
sp_lns <- SpatialLines(list(lns))
str(sp_lns)
```
 
---
 
## III. Add attributes (_Optional_:) 

Add attribute data, which will turn your `Spatial*` object into a `Spatial*DataFrame` object. ID fields are here required to match the data frame row names.

---

Create a simple `SpatialLinesDataframe`:

```{r}
dfr <- data.frame(id = "a", use = "road", cars_per_hour = 10) # note how we use the ID from above!
sp_lns_dfr <- SpatialLinesDataFrame(sp_lns, dfr, match.ID = "id")
str(sp_lns_dfr)
```

---

## Commonly used spatial methods

function | and what it does
------------ | ------------------------------------------------------
`bbox()` | returns the bounding box coordinates
`proj4string()` | sets or retrieves projection attributes using the CRS object.
`CRS()` | creates an object of class of coordinate reference system arguments
`spplot()` | plots all the attributes
`coordinates()` | set or retrieve the spatial coordinates.
`over(a, b)` | used to retrieve the polygon or grid indices on a set of points
`spsample()` | sampling of spatial points within the spatial extent of objects

---

# `sf` package

---

[`sf`](https://cran.r-project.org/package=sf) implements a formal standard called ["Simple Features"](https://en.wikipedia.org/wiki/Simple_Features) that specifies a storage and access model of spatial geometries (point, line, polygon). A feature geometry is called simple when it consists of points connected by straight line pieces, and does not intersect itself. 

This standard has been adopted widely, not only by spatial databases such as PostGIS, but also more recent standards such as GeoJSON. 

## How simple features in R are organized

Package `sf` represents simple features as native R objects.

All spatial functions and methods in `sf` prefixed by `st_` (refering to _spatial and temporal_)

Simple features are implemented as
R native data, using simple data structures (S3 classes, lists,
matrix, vector).  

---

Attributes typically stored in `data.frame` objects (or the
very similar `tbl_df`), we will also store feature geometries in
a `data.frame` column. Since geometries are not single-valued,
they are put in a list-column, a list of length equal to the number
of records in the `data.frame`, with each list element holding the simple
feature geometry of that feature.  The three classes used to
represent simple features are:

## Components

* `sf`, the table (`data.frame`) with feature attributes and feature geometries, which contains
* `sfc`, the list-column with the geometries for each feature (record), which is composed of
* `sfg`, the feature geometry of an individual simple feature.
---


If you work with PostGis or GeoJSON you may have come across the [WKT (well-known text)](https://en.wikipedia.org/wiki/Well-known_text) format, for example like these: 

    `POINT (30 10)
    LINESTRING (30 10, 10 30, 40 40)
    POLYGON ((30 10, 40 40, 20 40, 10 20, 30 10))`

`sf` implements this standard natively in R. Data are structured and conceptualized very differently from the `sp` approach.

---

In `sf` spatial objects are stored as a simple data frame with a special column that contains the information for the geographic coordinates. That special column is a list with the same length as the number of rows in the data frame. Each of the individual list elements then can be of any length needed to hold the coordinates that correspond to an individual feature.  

---

## I. Create geometric objects (topology)  

Geometric objects (simple features) can be created from a numeric vector, matrix or a list with the coordinates. They are called `sfg` objects for Simple Feature Geometry.

See here for an example of how a LINESTRING `sfg` object is created:
```{r}
lnstr_sfg <- st_linestring(matrix(runif(6), ncol=2)) 
class(lnstr_sfg)
```

---

## II. Combine all individual single feature objects for the special column. 

In order to work our way towards a data frame for all features we create what is called an `sfc` object with all individual features, which stands for Simple Feature Collection. The `sfc` object also holds the bounding box and the projection information.

---

See here for an example of how a `sfc` object is created:
```{r}
(lnstr_sfc <- st_sfc(lnstr_sfg)) # just one feature here
class(lnstr_sfc) 
```

---

## III. Add attributes. 

We now combine the dataframe with the attributes and the simple feature collection.
See here how its done.
```{r}
(lnstr_sf <- st_sf(dfr , lnstr_sfc))
class(lnstr_sf)
```

There are many methods available in the `sf` package, to find out use `methods(class="sp")`

---

## sf Highlights

* provides **fast** I/O, particularly relevant for large files 
* directly reads from and writes to spatial **databases** such as PostGIS
* compatibile with the *tidyverse*
* stay tuned for a new `ggplot` release that will be able to read and plot the `sf` format without the need of conversion to a data frame, like the `sp` format

---

`sp` and `sf` are _not_ only formats for spatial objects. Other spatial packages may use their own class definitions for spatial data (for example `spatstat`). Usuallly you can find functions that convert `sp` and increasingly `sf` objects to and from these formats.

---

```{r}
world_sp = as(world, "Spatial")
world_sf = st_as_sf(world_sp)
str(world_sp)
```

---

```{r}
str(world)
```

---



- The structures in the **sp** packages are more complicated - `str(world_sf)` vs `str(world_sp)`
  
- Moreover, many of the **sp** functions are not "pipeable" (it's hard to combine them with the **tidyverse**)

```{r, eval = F}
world_sp %>% 
  filter(name_long == "England")
```

`Error in UseMethod("filter_") : 
  no applicable method for 'filter_' applied to an object of class "c('SpatialPolygonsDataFrame', 'SpatialPolygons', 'Spatial', 'SpatialPolygonsNULL', 'SpatialVector')"`

---

## Reading and writing spatial data

```{r}
vector_filepath = system.file("shapes/world.gpkg", package = "spData")
vector_filepath
world = st_read(vector_filepath)
```

Counterpart to `st_read()` is the `st_write` function, e.g. `st_write(world, 'data/new_world.gpkg')`. A full list of supported formats could be found using `sf::st_drivers()`.

---

## Structure of the sf objects 

```{r, eval = FALSE}
world
```

```{r, echo = FALSE}
print(world, n=3)
```

```{r}
class(world)
```

---

## Structure of the sf objects 

```{r, eval=FALSE}
world$name_long
```

```{r, echo=FALSE}
world$name_long[1:3]
```

```{r, eval=FALSE}
world$geom
```

```{r, echo=FALSE}
print(world$geom, n = 3)
```


## Non-spatial operations on the sf objects

```{r, warning=FALSE}
world %>% 
  left_join(worldbank_df, by = "iso_a2") %>%
  select(name_long, pop, pop_growth, area_km2) %>%
  arrange(area_km2) %>% 
  mutate(pop_density = pop/area_km2) %>%
  rename(population = pop)
```

---

## Non-spatial operations

```{r}
world_cont = world %>% 
        group_by(continent) %>% 
        summarize(pop_sum = sum(pop, na.rm = TRUE))
```

```{r, echo=FALSE}
print(world_cont, n = 1)
```

---

The `st_set_geometry` function can be used to remove the geometry column:

```{r}
world_df =st_set_geometry(world_cont, NULL)
class(world_df)
```

---

## Spatial operations

It's a big topic which includes:

- Spatial subsetting
- Spatial joining/aggregation
- Topological relations
- Distances
- Spatial geometry modification
- Raster operations (map algebra)

See [Chapter 4](http://robinlovelace.net/geocompr/spatial-data-operations.html#spatial-operations-on-raster-data) of *Geocomputation with R* 

## CRS

Transform (warp) to a different projection:
```{r}
na_2163 = world %>%
  filter(continent == "North America") %>% 
  st_transform(2163) #US National Atlas Equal Area
st_crs(na_2163)
```

---

## Compare projections

```{r}
na = world %>%
  filter(continent == "North America") 
par(mfrow = c(1, 2), mar=c(0,0,0,0))
plot(na[0]);plot(na_2163[0])
```

---

##  The `world` dataset

```{r}
plot(na_2163)
```

---

## Spatial operations

```{r, warning = FALSE, message = FALSE, fig.height = 4}
canada = na_2163 %>% 
  filter(name_long=="Canada")

canada_buffer=canada%>%
  st_buffer(500000) %>%
  filter()

plot(select(canada,geom))
plot(select(canada_buffer,geom),add=T,border="red")

```


## Basic maps

- Basic maps of `sf` objects can be quickly created using the `plot()` function:

```{r}
plot(world[0])
```

```{r}
plot(world["pop"])
```

---

## [tmap](https://cran.r-project.org/web/packages/tmap/vignettes/tmap-nutshell.html): Thematic Maps
Syntax for creating plots is similar to that of ggplot2. 
The add-on package `tmaptools` contains tool functions for reading and processing shape files.

---

```{r, fig.align='center', fig.height=4, message=FALSE}
library(tmap)
tmap_mode("plot") #check the "view"
tm_shape(world, projection="wintri") +
        tm_polygons("lifeExp", title=c("Life expactancy"),
          style="pretty", palette="RdYlGn", auto.palette.mapping=FALSE) +
        tm_style_grey()
```

---

## leaflet
```{r}
library(leaflet)
library(widgetframe)
```

```{r, eval=F}
leaflet(world) %>%
        addTiles() %>%
        addPolygons(color = "#444444", weight = 1, fillOpacity = 0.5,
                    fillColor = ~colorQuantile("YlOrRd", lifeExp)(lifeExp),
                    popup = paste(round(world$lifeExp, 2)))
```

```{r, echo=T, message=T,results='asis'}
l = leaflet(world) %>%
  addTiles() %>%
  addPolygons(color = "#444444", weight = 1, fillOpacity = 0.5, 
              fillColor = ~colorQuantile("YlOrRd", lifeExp)(lifeExp),
              popup = paste(round(world$lifeExp, 2)))
saveWidget(l, file = "leaflet.html",selfcontained = T)
```
<iframe src="leaflet.html" width=100% height=400px></iframe>
---

## Raster data in the tidyverse

Raster data is not yet closely connected to the **tidyverse**, however:

- Some functions from the **raster** package works well in `pipes`
- You can convert vector data to the `Spatial*` form using `as(my_vector, "Spatial")`before working on raster-vector interactions
- There are some initial efforts to bring raster data closer to the **tidyverse**, such as [tabularaster](https://github.com/hypertidy/tabularaster), [sfraster](https://github.com/mdsumner/sfraster), or [fasterize](https://github.com/ecohealthalliance/fasterize)
- The development of the [stars](https://github.com/r-spatial/stars), package focused on multidimensional, large datasets should start soon. It will allow pipe-based workflows

---

## Sources

- Slides adapted from:
    - "Robin Lovelace and Jakub Nowosad" draft book [_Geocomputation with R_ (to be published in 2018)](http://robinlovelace.net/geocompr/). Source code at https://github.com/robinlovelace/geocompr.
    - [Claudia Engel's spatial analysis workshop](https://github.com/cengel/rspatial/blob/master/2_spDataTypes.Rmd)

